---
editor_options: 
  markdown: 
    wrap: 72
---

#load libraries

```{r}
#install.packages("scales")
library(scales)
#install.packages('ggrepel')
library(ggrepel)
#install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
#install.packages("lares")
library(lares)
#install.packages("stringi")
library(stringi) 
#install.packages("ggplot2")
library(ggplot2)
#install.packages("plyr")
library(plyr)
#install.packages("dplyr")
library(dplyr)
#install.packages("tidyr")
library(tidyr)
#install.packages("purrr")
library(purrr)
#install.packages("readr")
library(readr)
#install.packages("forcats")
library(forcats)
#install.packages("lattice")
library(lattice)
#install.packages("caret")
library(caret)
#install.packages("gridExtra")
library(gridExtra)
#install.packages("corrplot")
library(corrplot)
#install.packages("survival")
library(survival)
#install.packages("Hmisc")
library(Hmisc)
#install.packages("knitr")
library(knitr)
#install.packages("gridExtra")
library(gridExtra)
#install.packages("randomForest")
library(randomForest) 
#install.packages("ModelMetrics")
library(ModelMetrics)

```

#load data

```{r}
mydata <- read.csv("C:/Users/HP/Downloads/property-sales.csv", header=TRUE) 
summary(mydata)
```

#check columns names and record

```{r}
names(mydata)
```

```{r}
dim(mydata)
```

#structure of data

```{r}
str(mydata)
```

#check null values

```{r}
colSums(is.na(mydata))
```

#EDA

Find SalePrice distribution as we need to find all other features in
correlation with SalePrice

```{r}
ggplot(data=mydata[!is.na(mydata$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000),labels = comma)
```

, the sale prices are right skewed. This was expected as few people can
afford very expensive houses.

```{r}
```

```{r}
sapply(mydata, class)
```

```{r}
numericVars <- which(sapply(mydata, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')
```

Take all the numeric data

```{r}
numdata = mydata[ ,sapply(mydata, is.integer)]
#numdata = subset(numdata, select = -c(SalePrice) )
numdata
```

```{r}
all_numVar <- mydata[, numericVars]
cor_numVar <- cor(numdata, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```

```{r}
ggplot(data=mydata[!is.na(mydata$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot(col='blue') + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

```{r}
ggplot(data=mydata[!is.na(mydata$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma) +
        geom_text_repel(aes(label = ifelse(mydata$GrLivArea[!is.na(mydata$SalePrice)]>4500, rownames(mydata), '')))
```

We will remove outliers (524,1299) as they can affect model accuracy.

```{r}
mydata[c(524, 1299), c('SalePrice', 'GrLivArea', 'OverallQual')]
```

Encoding Categorical Data

1.MS Zoning

```{r}
mydata$MSZoning[is.na(mydata$MSZoning)] <- names(sort(-table(mydata$MSZoning)))[1]
mydata$MSZoning <- as.factor(mydata$MSZoning)
table(mydata$MSZoning)
```

```{r}
sum(table(mydata$MSZoning))
```

2.  BldgType

```{r}
mydata$BldgType[is.na(mydata$BldgType)] <- names(sort(-table(mydata$BldgType)))[1]
mydata$BldgType <- as.factor(mydata$BldgType)
table(mydata$BldgType)
```

```{r}
sum(table(mydata$BldgType))
```

3.House Style

```{r}
mydata$HouseStyle[is.na(mydata$HouseStyle)] <- names(sort(-table(mydata$HouseStyle)))[1]
mydata$HouseStyle <- as.factor(mydata$HouseStyle)
table(mydata$HouseStyle)
```

```{r}
sum(table(mydata$HouseStyle))
```

4.Central Air

```{r}
mydata$CentralAir[is.na(mydata$CentralAir)] <- names(sort(-table(mydata$CentralAir)))[1]
mydata$CentralAir <- as.factor(mydata$CentralAir)
table(mydata$CentralAir)
```

5.KitchenQual

```{r}
mydata$KitchenQual[is.na(mydata$KitchenQual)] <- names(sort(-table(mydata$KitchenQual)))[1]
mydata$KitchenQual <- as.factor(mydata$KitchenQual)
table(mydata$KitchenQual)
```

6.SaleCondition

```{r}
mydata$SaleCondition[is.na(mydata$SaleCondition)] <- names(sort(-table(mydata$SaleCondition)))[1]
mydata$SaleCondition <- as.factor(mydata$SaleCondition)
table(mydata$SaleCondition)
```

7.Fireplace

```{r}
mydata$Fireplace[is.na(mydata$Fireplace)] <- names(sort(-table(mydata$Fireplace)))[1]
mydata$Fireplace <- as.factor(mydata$Fireplace)
table(mydata$Fireplace)
```

```{r}
dim(mydata)
```

```{r}
numericVars <- which(sapply(mydata, is.integer)) #index vector numeric variables
factorVars <- which(sapply(mydata, is.factor)) #index vector factor variables
cat('There are', length(numericVars), 'numeric variables, and', length(factorVars), 'categoric variables')
```

**I wanted to get an overview of the most important variables, including
the categorical variables, before moving on to visualization, despite
the correlations giving a good overview of the numeric variables.**

```{r}
set.seed(2018)
quick_RF <- randomForest(x=mydata[1:1460,-17], y=mydata$SalePrice[1:1460], ntree=200,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[2:17,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")
```

Plotting most important categorical feature

The first graph shows the median SalePrice by MSZoning. The frequency
(number of houses) of each MS Zone in the data set is shown in the
labels.

The second graph below shows the frequencies across all data.

```{r}
n1 <- ggplot(mydata[!is.na(mydata$SalePrice),], aes(x=KitchenQual, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
n2 <- ggplot(data=mydata, aes(x=KitchenQual)) +
        geom_histogram(stat='count')+
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3)+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(n1,n2)
```

### Categorical feature 2

The first graph shows the median SalePrice by FirePlace. The frequency
(number of houses) with Fireplace in the data set is shown in the
labels.

The second graph below shows the frequencies across all data.

```{r}
ms1 <- ggplot(mydata[!is.na(mydata$SalePrice),], aes(x=HouseStyle, y=SalePrice)) +
        geom_bar(stat='summary', fun.y = "median", fill='blue') +
        theme(axis.text.x = element_text( hjust = 1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=50000), labels = comma) +
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        geom_hline(yintercept=163000, linetype="dashed", color = "red") #dashed line is median SalePrice
ms2 <- ggplot(data=mydata, aes(x=HouseStyle)) +
        geom_histogram(stat='count')+
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3) +
        theme(axis.text.x = element_text( hjust = 1))
grid.arrange(ms1,ms2)
```

There are 2 bathroom variables. Individually, these variables are not
very important. However, I assume that I if I add them up into one
predictor, this predictor is likely to become a strong one.

```{r}
mydata$TotBathrooms <- mydata$FullBath + (mydata$HalfBath*0.5) 
```

As you can see in the first graph, there now seems to be a clear
correlation . The frequency distribution of Bathrooms in all data is
shown in the second graph.

```{r}
tb1 <- ggplot(data=mydata[!is.na(mydata$SalePrice),], aes(x=as.factor(TotBathrooms), y=SalePrice))+
        geom_point(col='blue') + geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
tb2 <- ggplot(data=mydata, aes(x=as.factor(TotBathrooms))) +
        geom_histogram(stat='count')
grid.arrange(tb1, tb2)
```

```{r}
mydata
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}
```

```{r}


```

```{r}

```

```{r}

```

```{r}

```
